An Extensible Approach to Verification of Embedded Network Systems*

### 2014-04-21 tech-project/paper.tex
paper updated

In this paper we present a flexible means of specifying and verifying the correctness of software for embedded network systems. Our approach uses RESOLVE, an imperative, component based programming and mathematical specification language, to verify the functional correctness of TinyOS applications. In doing so, we enrich the work originally presented in \cite{regula:2010} with the following: A model view controller (MVC) based implementation of a RESOLVE to C translator, a dynamic memory allocation scheme tailored towards embedded systems running the code generated by our tool, and the addition of a new language keyword that enables users to pair custom RESOLVE specifications with `externally' implemented (non-native RESOLVE) realizations. We demonstrate these additions on an LED (Light emitting diode) driver that showcases recent mathematical developments, as well as formal verification of two enhancements that we demonstrate running on a Telos mote.
\end{abstract}

@@

The concept provides two operations. The first of the two, \texttt{Set}, takes as a parameter an instance of an LED strip \texttt{L}, a boolean \texttt{b}, and an integer \texttt{i}. The operation \texttt{requires} that \texttt{i} falls within the length of the strip, and \texttt{ensures} upon completion that:
\begin{enumerate}
\item The length of the outgoing strip \texttt{L}, is the same length as the incoming strip, \texttt{\#L} (where \texttt{\#} signifies its incoming state).
\item The LED in position \texttt{i} of \texttt{L} is set to boolean parameter \texttt{b}. 
\end{enumerate}
Operation \texttt{Status}, which simply returns the state of the LED at position \texttt{i}, is specified similarly.

@@

RESOLVE also allows users to extend the functionality provided by the base concept through \textit{enhancements} -- a form of specification inheritance. The enhancement we provide here is  \texttt{Toggling\_Capability}, which gives users of the LED component the ability to toggle a specific LED on or off.

@@

The specification of \texttt{Toggle} states that upon termination, the LED located at position \texttt{i} in \texttt{L} is the complement of that same location in the incoming LED, \texttt{\#L}.

@@

Even realizations of enhancements are neutral since they are only permitted to call operations provided by the base concept -- thus requiring no knowledge of how the called operation is actually implemented.

@@

In this section we verify our realization of \texttt{Toggling\_Capability}. The first step in this process is to generate verification conditions (VCs) for \texttt{Toggling\_Realiz}, which, if proven, will establish the correctness of that particular realization\footnote{Note: VCs themselves are typically generated from specific lines of a realization to ensure that the overall content of a realization is consistent with its specification.}.

\begin{figure}[!htb]
\centering
\begin{tabular}{lccc}
	\toprule
	Condition \# & Time (ms)	& Steps	& Search \\
	\midrule
	VC 0\_1	& 4426	& 5	& 0	\\
	VC 0\_2	& 5039	& 5	& 0	\\
	VC 0\_3	& 6324	& 5	& 0	\\
	\bottomrule
\end{tabular}
\end{figure}

The results of verification are summarized Figure \ref{fig:results}. As shown, the correctness of \texttt{Toggle} is straightforward enough for the prover to establish automatically, however \texttt{Set\_First\_And\_Last}'s VC 0\_5 failed to be dispatched.

\begin{figure}[!htb]
\begin{verbatim}
Goal:
Element_At((|L'|  - |L'|), L') = b

Given:
    ...
1. |L| = Strip_Length and
2. |L|  > 1 and
3. |L''| = |L| and
4. Element_At(0, L'') = b
5. |L'| = |L''| and
6. Element_At((|L''|  - 1), L') = b)
\end{verbatim}
\caption{A problematic VC}
\label{fig:problematicVC}
\end{figure}

A cursory examination of the unproven VC, shown in Figure \ref{fig:problematicVC}, reveals that prover was not able to employ the seemingly relevant \texttt{given \#4} since \texttt{L"}, could not be considered (rightfully so) the same as \texttt{L'} due to the changing boolean value. There are several means of correcting this: The first involves amending the \texttt{ensures} clause of operation \texttt{Set} (which is used in \texttt{Set First And Last}) to intuitively say that ``the content of \textit{outgoing} \texttt{L} is equivalent to the that of the incoming \texttt{L} with the exception of position \texttt{i}. The second route -- perhaps the less preferable of the two -- involves adding new theorems to \texttt{String\_Theory} which state this.

FIGURE OUT AN ELEGANT WAY TO CONCLUDE THIS SECTION...

\subsection{Facilities}
\label{sec:facilities}
The most straightforward means of utilizing our previously defined Led strip component and its verified enhancements is a via \textit{facility} module. Figure \# depicts a simple

>>>>

In this paper we present a flexible means of specifying and verifying the correctness of software for embedded network systems. Our approach uses RESOLVE, an imperative, component based programming and mathematical specification language, to verify the functional correctness of TinyOS applications. In doing so, we enrich the work originally presented in \cite{regula:2010} with the following: A model view controller (MVC) based implementation of a RESOLVE to C translator, a dynamic memory allocation scheme tailored towards embedded systems running the code generated by our tool, and the addition of a new language keyword that enables users to pair custom RESOLVE specifications with `externally' implemented (non-native RESOLVE) realizations. We demonstrate these additions on an LED (Light emitting diode) driver that showcases recent mathematical developments, as well as formal verification of a toggling capability enhancement that we demonstrate running on a Telos mote.
\end{abstract}

@@

The concept provides two operations. The first operation, \texttt{Set}, takes as a parameter an instance of an LED strip \texttt{L}, a boolean \texttt{b}, and an integer \texttt{i}. The operation \texttt{requires} that \texttt{i} falls within the length of the strip, and \texttt{ensures} upon completion of the operation that the length of the outgoing strip \texttt{L} is the same length as the incoming strip, \texttt{\#L}, and that the LED in position \texttt{i} of \texttt{L} is set to boolean \texttt{b}. The \texttt{Status} operation is specified similarly.

@@

RESOLVE also allows users to extend the functionality provided by the base concept through \textit{enhancements} -- a form of specification inheritance. The enhancement we provide here, \texttt{Toggling\_Capability}, allows users to flip a specific LED to its complement.

@@

The enhancement adds a single operation, \texttt{Toggle}, which states that upon termination, the LED located at position \texttt{i} in \texttt{L} is the complement of that same location in the incoming LED, \texttt{\#L}.

@@

Even enhancement realizations are neutral since only the operations called are those provided by the base concept -- and thus require no knowledge of how the called operation is implemented.

In this section we give a short overview of the verification results of \texttt{Toggling\_Realiz}. The first step in proving this particular realization correct is to generate verification conditions (VCs), which, if proven, will establish the correctness of this particular realization\footnote{Note: VCs themselves are typically generated from specific lines of a realization to ensure that the overall content of a realization is consistent with its specification.}.

\begin{figure}[!htb]
\centering
\begin{tabular}{lccc}
	\toprule
	Condition \# & Time (ms)	& Steps	& Search \\
	\midrule
	VC 0\_1	& 4426	& 5	& 0	\\
	VC 0\_2	& 5039	& 5	& 0	\\
	VC 0\_3	& 6324	& 6	& 0	\\
	\bottomrule
\end{tabular}
\caption{Verification results for \texttt{Toggle} (left)}
\label{fig:results}
\end{figure}

The results of verification are summarized Figure \ref{fig:results}. Using the RESOLVE compiler's integrated prover, we are able to mechanically and automatically dispatch all VCs generated by operation \texttt{Toggle}, thus verifying its correctness. In terms of proof difficulty, given the number of steps and time taken to establish each, leads us to conclude that the VCs generated were of a straightforward variety. Though readers interested in a more detailed discussion of how exactly the prover goes about transforming and dispatching VCs are encouraged to refer to \cite{smith:2013}.

\subsection{Facilities}
\label{sec:facilities}

With our formally specified LED strip component in place -- and a verified enhancement on this component -- we now present a sample embedded application that iteratively toggles all lights in a single strip.

Shown below is a RESOLVE facility module that implements the client logic 
\begin{verbatim}
Facility LED_Telos_Demo;
    uses Std_Clock_Fac;

    Facility Leds_Fac is LED_Template(3)
        externally realized by Std_Led_Realiz
     enhanced by Toggling_Capability
        realized by Toggling_Realiz;

    Operation Main(); Procedure
        Var I1, I2, I3 : Integer;
        Var Loop : Boolean;
        Var L : Led;

        I1 := 1; I2 := 2; I3 := 3;

        Loop := True();
        While(Loop)
            changing Loop;
            maintaining ...
        do
            Leds_Fac.Toggle(L, I1);
            Std_Clock_Fac.Wait_500_Milli_Seconds();

            Leds_Fac.Toggle(L, I2);
            Std_Clock_Fac.Wait_500_Milli_Seconds();

            Leds_Fac.Toggle(L, I3);
            Std_Clock_Fac.Wait_500_Milli_Seconds();
        end;
    end Main;

end LED_Telos_Demo;
\end{verbatim}

Prior to manipulating the LED component developed in the previous sections, we must first pair its specification with a realization. This is accomplished via a facility declaration:

@@
